\ProjectName{}: Data Imputation as a Query Optimization

### 2016-12-14 paper/algorithm.tex
paper: Add discussion of time complexity of tree build

\todo{is it necessary to add a fudge factor to the loss calculation, for generality, and set
it at 1.0 for our paper?}

>>>>

To compute $\text{Time}(Q)$, we retain the heuristics used by SimpleDB for the time
complexity of the typical relational operators, and add new formulations for \textit{Drop}
and \textit{Impute}. \textit{Drop} is a special case of a sequential scan, and its time
complexity can be expressed as a function of the number of heap pages read and the IO cost
per page.

Evaluation of the time complexity for \textit{Impute} is more thorny, as the properties of
different decision tree building algorithms can vary significantly, the underlying IO cost
is not easily extracted (especially in the case that the entire stream of tuples does not
fit in memory), and the CPU cost dominates (in contrast to the other operators, which may
not even consider CPU cost explicitly). For example, \cite{martin1995time} find that the
time complexity of the \textit{build tree} algorithm for one commonly-used class of
decision trees is a function of the number of classes, several overhead constants, the
parameterization of the partition and heuristic functions, and the \textit{arity} (the
number of subsets considered for each split). Indeed, the \textit{build tree} phase often
does not even dominate computation, as post-processing steps like pruning, which are vital
in achieving good performance, can have cost exponential in the height of the tree. In
practice, we find that a simple parameterization can be acceptable (yield intelligent query
plans), though the range of $\alpha$ that does promote tradeoffs is condensed.
